---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2023-03-03"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSQLite)
library(tidyverse)
library(caret)
library(magrittr)
library(glue)
library(cowplot)
library(modelr)
```
## Reading in the data
Read in the data, and set the data types to the correct types
Drop index number from the pandas dataframe and NCBI.tax.ID
```{r}
# Connect to the DB
conn <- dbConnect(SQLite(),"../../../s16.sqlite")
# List of all the tables
dbListTables(conn)

D_tmp <- dbGetQuery(conn, "SELECT * FROM ribdif_bacdive_joined")
D_tmp <- tibble(D_tmp, .name_repair ="universal")

D_tmp <- mutate(D_tmp, across(antibiotics:PH.range, factor))
D_tmp <- select(D_tmp, !c(NCBI.tax.ID,strain.designation))

# Chainging AR with no annotation to PNR
D_tmp <- mutate(D_tmp, antibiotics = ifelse(is.na(antibiotics), "PNR", "R"))
```



# Splitting up data

```{r}
set.seed(25022023)
# Adding ID as a column
D_tmp %<>% mutate(ID = row_number(species))
# Randomly selecting the training/exploration data with seed set
D <- D_tmp %>% slice_sample(prop = 0.7)
# Assigning the rest of the data to the test dataset
D_test <- anti_join(D_tmp, D, by = "ID")

# Checking if its correctly split up
percent_in_test <- nrow(D_test)/(nrow(D)+nrow(D_test))
percent_in_test

```

## Modeling
The goal of this part is to build a model which takes into consideration 
n16 and taxonomic relationships as it seems they might have a big impact.
We could either:
1) Remove all positions with n16 = 0, as they are not going to include any information about 
the relationship between bacterial ecology and div. Here the intercept could now be set to (0,0) or not
2) Fit a model with a varying intercept
3) Fit a model with 0,0 as intercept as described in the next paragraph
Below we can see div against n16.
The main takeway is that we have a lot small values and a few large, therefore there is an arguemt
for applying a transformation to both axis.

```{r}
ggplot(D,aes(x=n16, y=div)) + 
  geom_point(size = 0.2) +
  geom_smooth()
```

## Transformations

```{r}
Dt <- D %>% 
  mutate(Tn16=log(n16), Tdiv=log1p(div)) 

# Plotting the different transformations
p1 <- ggplot(Dt,aes(x=Tn16, y=Tdiv)) + 
  geom_point(size=0.2) +
  geom_smooth()

p2 <- ggplot(Dt,aes(x=n16, y=Tdiv)) + 
  geom_point(size=0.2) +
  geom_smooth()

p3 <- ggplot(Dt,aes(x=n16, y=div)) + 
  geom_point(size=0.2) +
  geom_smooth()

plot_grid(p1,p2,p3,labels ="auto")

```
## Looking at residuals for each
```{r}
a <- lm(Tdiv ~ Tn16  ,Dt)
b <- lm(Tdiv ~ n16  ,Dt)
c <- lm(div ~ n16 ,Dt)

```
a)
```{r}
par(mfrow=c(2,2))
plot(a)
```
b)
```{r}
par(mfrow=c(2,2))
plot(b)
```
c)
```{r}
par(mfrow=c(2,2))
plot(c)
```

Based on this im going to go with the transformation of both sides. Since these transformations decrease the leverage of larger numbers


### Taxonomic information in the model
Lets try and add taxonomic information to the mdoel
Lets first visualize the phylums with over 20 entries
Here it seems that there could be some gain in including phylum in the model, as it seems to have an effect

```{r}
# Plotting for different phylum
Dt %>% 
  group_by(phylum) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 50) %>% 
  ggplot(aes(x=Tn16, y=Tdiv)) +
    geom_point(size=0.5, alpha=0.6) +
    theme(legend.position="none") +
    facet_wrap(~phylum) +
    geom_smooth(method=lm, formula = "y~x") +
    ggtitle("By phylum")


```
Lets also have a look for class, 
```{r}
library(magrittr)
# Plotting for different orders
Dt %>% 
  group_by(class) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20) %>% 
  ggplot(aes(x=Tn16, y=Tdiv)) +
    geom_point(size=0.5, alpha=0.6) +
    theme(legend.position="none") +
    facet_wrap(~class) +
    geom_smooth(method=lm, formula = "y~x+0") +
    ggtitle("By class")

```
Here it's hard to see how much information we lose by just including phylum instead of order. Therefore lets
try and remove the effect of phylum by plotting the residuals of a simple model

```{r}
fitTaxPhylum <- lm(Tdiv ~ 0 + Tn16 + Tn16:factor(phylum)  ,Dt)
summary(fitTaxPhylum)
res <- Dt %>% 
  add_residuals(fitTaxPhylum)

res %>% 
  group_by(class) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20) %>% 
  ggplot(aes(x=Tn16, y=resid)) +
    geom_point(size=0.5, alpha=0.6) +
    theme(legend.position="none") +
    facet_wrap(~class) +
    geom_smooth(method=lm, formula = "y~x+0") +
    geom_ref_line(h=0, col = "red", size = 0.1) +
    ggtitle("Residuals vs n16 By class")


```
The residuals seems ok distributed. It seems to make sense stay at the phylum level just based on this
But checking the amount of entries in both it seems that they are about the same.
So this effect could be due to each phylum just having one class.
```{r}
print("class:")
res %>% 
  group_by(class) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20) %>% 
  nrow()
print("phylum:")
res %>% 
  group_by(phylum) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20) %>% 
  nrow()

```
Lets check
```{r}
res %>% 
  group_by(phylum) %>% 
  reframe(uClass = unique(class)) %>% 
  ggplot() + geom_bar(aes(x=phylum)) +coord_flip()


```
This seems to be the case
We could also do the same for order.
Here there is more varibilty. But it will argue that we get closer to just predicting the datapoints directly instead of 
the tendency. Therefore i am going to just keep the model with including the phylum level

```{r}
fitTaxPhylum <- lm(Tdiv ~ 0 + Tn16 + Tn16:phylum ,Dt)
summary(fitTaxPhylum)
res <- Dt %>% 
  add_residuals(fitTaxPhylum)

res %>% 
  group_by(order) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20) %>% 
  ggplot(aes(x=Tn16, y=resid)) +
    geom_point(size=0.5, alpha=0.6) +
    theme(legend.position="none") +
    facet_wrap(~order) +
    geom_smooth(method=lm, formula = "y~x+0") +
    geom_ref_line(h=0, col = "red", size = 0.1) +
    ggtitle("Residuals vs n16 By order")
```

## Looking at tax + n16 model


```{r}
Dt <- Dt %>% 
  group_by(phylum) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20)
# Lets add it to the model
Dt
fitTaxPhylum <- lm(Tdiv ~ Tn16 + phylum + Tn16:phylum ,Dt)
summary(fitTaxPhylum)
res <- Dt %>% 
  add_residuals(fitTaxPhylum)

res %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res %>% 
  ggplot(aes(x=exp(Tn16) ,y=resid)) +
  geom_hex(bins=50) +
  geom_ref_line(h=0, col = "red", size = 1) +
  geom_smooth()

par(mfrow=c(2,2))
plot(fitTaxPhylum )

# We have on with a very large amount of n16
#Dsub %>% filter(n16>20)

#I checked and its also high here
#https://www.arb-silva.de/search/
#Tumebacillus avium
# we observe more var at the start in the res since they are predicting wrong

```
We can observe that we tend to overestimate the div on genera with larger amount of #16s. 
And we tend to underestimate div for genera with samller amounts of #16s.
Therefore we still have some unexplained variance in the model

Lets also have a look for order

```{r}
Dt <- Dt %>% 
  filter(n16 > 1)
# Lets add it to the model
Dt
fitTaxOrder <- lm(Tdiv ~ Tn16 + Tn16:order ,Dt)
res <- Dt %>% 
  add_residuals(fitTaxOrder)

res %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res %>% 
  ggplot(aes(x=exp(Tn16) ,y=resid)) +
  geom_hex(bins=50) +
  geom_ref_line(h=0, col = "red", size = 1) +
  geom_smooth()

par(mfrow=c(2,2))
plot(fitTaxPhylum )

```
While some of the plots look ok,
we can see that most of the orders have few entries (~half having below 10 entries)

```{r}
# Under 10
print("under 10")
Dt %>% 
  group_by(order) %>% 
  summarise(n = n()) %>% 
  filter(n < 10) %>% 
  nrow

# Over or equal to 10
print("over or equal to 10")
Dt %>% 
  group_by(order) %>% 
  summarise(n = n()) %>% 
  filter(n >= 10) %>% 
  nrow

```

### SAMPLING
Lets try and get an idea about the effect of where it is samples from
```{r}

Dt <- D %>% 
  mutate(Tn16=log(n16), Tdiv=log1p(div)) 
Dt <- Dt %>% 
  group_by(phylum) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 20)

Denv <- Dt %>% mutate(aquaP = aquatic.counts/Total.samples , 
              animalP = animal.counts/Total.samples, 
              plantP = plant.counts/Total.samples, 
              soilP = soil.counts/Total.samples)

fitTaxPhylum <- lm(Tdiv ~ Tn16 + Tn16:phylum ,Denv)
res <- Denv %>% 
  add_residuals(fitTaxPhylum)
summary(fitTaxPhylum)


p1 <- ggplot(res, aes(x=aquaP, y=resid)) +
  geom_point()
p2 <- ggplot(res, aes(x=plantP, y=resid)) +
  geom_point()
p3 <- ggplot(res, aes(x=animalP, y=resid)) +
  geom_point()
p4 <- ggplot(res, aes(x=soilP, y=resid)) +
  geom_point()
plot_grid(p1, p2, p3, p4,labels ="auto")

```
It seems that there is no difference here

## Antibiotics, motility, PH, gramstain motility

```{r}




res <- Dt %>% 
  add_residuals(fitTaxPhylum)
summary(fitTaxPhylum)



p1 <- ggplot(res, aes(x=gram.stain, y=resid)) +
  geom_boxplot()
p2 <- ggplot(res, aes(x=PH.range, y=resid)) +
  geom_boxplot() +
  geom_point(size=0.2)
p3 <- ggplot(res, aes(x=motility, y=resid)) +
  geom_boxplot()+
  geom_point(size=0.2)
p4 <- ggplot(res, aes(x=antibiotics, y=resid)) +
  geom_boxplot() +
  geom_point(size=0.2)
plot_grid(p1, p2, p3, p4,labels ="auto")
```
Lets test antibiotics

```{r}
library(car)
# Updating model and running ancova on it
fit_ar <- update(fitTaxPhylum, . ~ . + factor(antibiotics) + Tn16:factor(antibiotics))
Anova(fit_ar)
drop1(fit_ar)
fit_ar2 <- update(fit_ar, .~. -Tn16:factor(antibiotics))
Anova(fit_ar2)

```
It seems there is no significant effect of antibiotics



#### Lets have a look at specific antibiotics

### Formatting daTA
Lets check for more specifc types of AR
First getting the subset of the data with AR resistence info about the Antibiotics
which target the 16s rRNA
```{r}
# Getting the ones which are actually targeting 16S 
# Reading them from ARtarget16s.csv
target16S <- read_csv2("../data/ARtarget16s.csv",show_col_types = FALSE,col_names = FALSE)
targetvector <- as.array(target16S$X1)
found_16S <- as.array(colnames(select(D_tmp,lincomycin:spiramycin.II)))
intersect <- intersect(targetvector,found_16S)
D_ar <- select(Dt, all_of(intersect), Tn16, Tdiv, phylum)
```
### Different types
#### Div
Now lets look at some plots
firstly for div
```{r}
library(patchwork)

plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(res)+
    geom_boxplot(aes(x=.data[[antibiotic]], y=resid)) +
    geom_point(aes(x=.data[[antibiotic]], y=resid), size = 0.5, alpha = 0.5)
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)
```

# Lets test it
```{r}

update(fitTaxPhylum, . ~ . + minocycline + Tn16:minocycline) %>% 
  Anova()

filter(D_ar, !is.na(minocycline ))

update(fitTaxPhylum, . ~ . + streptomycin + Tn16:streptomycin) %>% 
  Anova()

#update(fitTaxPhylum, . ~ . + tobramycin + Tn16:tobramycin) %>% 
#  Anova()


```




```{r}
plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(D_ar)+
    geom_bar(aes(x=.data[[antibiotic]],na.rm = TRUE))
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)
```





#### Lets have a look at interactions

