---
title: "Dataanalysis"
output: pdf_document
date: "2023-02-17"
---
## TODO
Test for TAX VS AR VS GENOME SIZE
mosaic plot?s

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library(xtable)
library(glue)
library(cowplot)
library(ggpubr)
library(patchwork)
```

## Reading in data
Read in the data, and set the data types to the correct types
Drop index number from the pandas dataframe and NCBI.tax.ID
```{r}

D_tmp <- read_csv("../data/dataset_joined.csv",show_col_types = FALSE)
D_tmp <- mutate(D_tmp, across(species:PH.range, factor))
D_tmp <- mutate(D_tmp, across(GC.content:total_genes, as.double))
D_tmp <- select(D_tmp, !c(NCBI.tax.ID,...1))

```

Split up the dataset in two, one with the specific antibiotic resistance information and another with the rest of the data

```{r}
# With ar
D_ar <- select(D_tmp, !antibiotics & species:spiramycin.II)
# Without ar
D <- select(D_tmp, !lincomycin:spiramycin.II)
```

## Summary Statistics

```{r}
sum <- summary(D)
sum
```
We have NA for the following columns: motility,gram.stain,growth,oxygen.tolerance,PH.range

## Looking at data distribution
### For different taxonomic scales
TODO: also make plot of just n species

```{r}
# Function to plot how often each tax is seen
taxplot <- function(D, tax, name,ylimit){
  plot <- D %>% 
    group_by({{tax}}) %>%
    summarise(n=n()) %>% 
    mutate(tax_ordered = fct_reorder({{tax}},desc(n))) %>% 
    ggplot() +
    geom_segment(aes(x=tax_ordered, xend = tax_ordered,y=0, yend=n)) +
    xlab(glue("taxonomic {name}")) + ylab("Number of entries") +
    theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
    ggtitle(glue("Entries pr. {name}")) +
    ylim(0,ylimit)
  return(plot)
}

# Plloting the different taxs
p1 <- taxplot(D, family, "family",350)
p2 <- taxplot(D, order, "order",350)
p3 <- taxplot(D, class, "class",2000)
p4 <- taxplot(D, phylum, "phylum",2000)
p5 <- taxplot(D, genus, "genus",350)
plot_grid(p5,p1, p2, p3, p4,labels ="auto")
#gridExtra::grid.arrange(p5,p1,p2,p3,p4)
```
The goal of this part is to figure out how to take into consideration that some strains are sequences a lot more than others.
I have already calculated the mean/mode of each value at the species level, but it seems that we still have an unequal distribution.
Furthermore it seems that nomatter how high the taxonomic scale we go, There is still an unequal distribution. Even though wee took into 
consideration that some species are sequences more we would still except an unequeal distribution. Based on this and the fact that information is lost
for each "step" on the taxonomic scale I have choosen to go with the speceis level
It is relevant as when we for example look at the distribution of numbers of 16s genes, more sequences genes will have a higher weight.

```{r}
taxboxplot <- function(D, tax, n16_var, name){
  D_new <- D %>% 
    group_by({{tax}}) %>%
    summarise(mean_value=mean({{n16_var}}))
  plot <- ggplot(D_new) +
    geom_histogram(aes(x=mean_value, y=after_stat(density))) +
    ggtitle(glue("density pr. {name}"))+
    xlab("mean n16") + ylab("density") +
    stat_function(fun = dnorm, col = "red", args = list(mean = mean(D_new$mean_value), sd = sd(D_new$mean_value)))
  return(plot)
}


# Plloting the different taxs
to_plot <- "n16"
p0 <- taxboxplot(D, species, .data[[to_plot]],"species")
p1 <- taxboxplot(D, genus, .data[[to_plot]],"genus")
p2 <- taxboxplot(D, family,.data[[to_plot]], "family")
p3 <- taxboxplot(D, order, .data[[to_plot]],"order")
p4 <- taxboxplot(D, class,.data[[to_plot]], "class")
p5 <- taxboxplot(D, phylum,.data[[to_plot]], "phylum")

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")


```
But the most important thing is the distribution. Here we can see that for species it looks a bit wierd, it looks better for genus
Lets have a look at the log transformed

```{r}
# Not working:: Error in if (!is.finite(log_num)) { : the condition has length > 1
L2log <- function(num){
  log_num = log2(num)
  print(log_num)
  if (!is.finite(log_num)){return(0)}
  else{
  return (log_num)}
}

functaxboxplot <- function(D, tax, n16_var, name,func){
  D_new <- D %>% 
    group_by({{tax}}) %>%
    summarise(mean_value=mean({{n16_var}}))
  plot <- ggplot(D_new) +
    geom_histogram(aes(x=func(mean_value), y=after_stat(density)),bins = 13) +
    ggtitle(glue("density pr. {name}"))+
    xlab("logn16") + ylab("density") +
    stat_function(fun = dnorm, col = "red", args = list(mean = mean(func(D_new$mean_value)), sd = sd(func(D_new$mean_value))))
  return(plot)
}


# Plloting the different taxs
to_plot <- "n16"
p0 <- functaxboxplot(D, (species), .data[[to_plot]],"species",log)
p1 <- functaxboxplot(D, (genus), .data[[to_plot]],"genus",log)
p2 <- functaxboxplot(D, (family),.data[[to_plot]], "family",log)
p3 <- functaxboxplot(D, (order), .data[[to_plot]],"order",log)
p4 <- functaxboxplot(D, (class),.data[[to_plot]], "class",log)
p5 <- functaxboxplot(D, (phylum),.data[[to_plot]], "phylum",log)

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")


```
Here the distributions look ok for species. And again to not lose information im going with species.s

#### D
Lets have a look at the distributions for D

```{r}

taxboxplot <- function(D, tax, n16_var, name){
  D_new <- D %>% 
    group_by({{tax}}) %>%
    summarise(mean_value=mean({{n16_var}}))
  plot <- ggplot(D_new) +
    geom_histogram(aes(x=mean_value, y=after_stat(density))) +
    ggtitle(glue("Entries pr. {name}"))+
    xlab("div") + ylab("density") +
    stat_function(fun = dnorm, col = "red", args = list(mean = mean(D_new$mean_value), sd = sd(D_new$mean_value)))
  return(plot)
}

to_plot <- "div"
p0 <- taxboxplot(D, species, .data[[to_plot]],"species")
p1 <- taxboxplot(D, genus, .data[[to_plot]],"genus")
p2 <- taxboxplot(D, family,.data[[to_plot]], "family")
p3 <- taxboxplot(D, order, .data[[to_plot]],"order")
p4 <- taxboxplot(D, class,.data[[to_plot]], "class")
p5 <- taxboxplot(D, phylum,.data[[to_plot]], "phylum")

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")


```
They do not look normal distributed.


## Looking at n16 and var

```{r}


functaxboxplot <- function(D, tax, n16_var, name,func){
  D_new <- D %>% 
    group_by({{tax}}) %>%
    summarise(mean_value=mean({{n16_var}}))
  plot <- ggplot(D_new) +
    geom_histogram(aes(x=func(mean_value), y=after_stat(density)),bins = 20) +
    ggtitle(glue("density pr. {name}"))+
    xlab("logdiv") + ylab("density") +
    stat_function(fun = dnorm, col = "red", args = list(mean = mean(func(D_new$mean_value)), sd = sd(func(D_new$mean_value))))
  return(plot)
}

# Plloting the different taxs
to_plot <- "div"
p0 <- functaxboxplot(D, (species), .data[[to_plot]],"species",log1p)
p1 <- functaxboxplot(D, (genus), .data[[to_plot]],"genus",log1p)
p2 <- functaxboxplot(D, (family),.data[[to_plot]], "family",log1p)
p3 <- functaxboxplot(D, (order), .data[[to_plot]],"order",log1p)
p4 <- functaxboxplot(D, (class),.data[[to_plot]], "class",log1p)
p5 <- functaxboxplot(D, (phylum),.data[[to_plot]], "phylum",log1p)

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")

```

Here we can see we have alot of divs at = 0,
Looking at not having them we get this
```{r}

# Plloting the different taxs
to_plot <- "div"
p0 <- functaxboxplot(D, (species), .data[[to_plot]],"species",log)
p1 <- functaxboxplot(D, (genus), .data[[to_plot]],"genus",log)
p2 <- functaxboxplot(D, (family),.data[[to_plot]], "family",log)
p3 <- functaxboxplot(D, (order), .data[[to_plot]],"order",log)
p4 <- functaxboxplot(D, (class),.data[[to_plot]], "class",log)
p5 <- functaxboxplot(D, (phylum),.data[[to_plot]], "phylum",log)

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")
```
There is an argument for not working with the var of 0 or we can try and remove all entries with n16=1 (since they always has zero)
It turns out n16=1 gives a ton with no var, n16 > 4 is more acurate. Alternativly you could have it for one speciefic n16 or take n16 into the model when modelling div
```{r}
# For n16 > 4
D_sub <- filter(D, n16 > 4)
to_plot <- "div"
p0 <- functaxboxplot(D_sub, (species), .data[[to_plot]],"species",log1p)
p1 <- functaxboxplot(D_sub, (genus), .data[[to_plot]],"genus",log1p)
p2 <- functaxboxplot(D_sub, (family),.data[[to_plot]], "family",log1p)
p3 <- functaxboxplot(D_sub, (order), .data[[to_plot]],"order",log1p)
p4 <- functaxboxplot(D_sub, (class),.data[[to_plot]], "class",log1p)
p5 <- functaxboxplot(D_sub, (phylum),.data[[to_plot]], "phylum",log1p)
plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")

# For n16 = 4
D_sub <- filter(D, n16 == 10)
to_plot <- "div"
p0 <- functaxboxplot(D_sub, (species), .data[[to_plot]],"species",log1p)
p1 <- functaxboxplot(D_sub, (genus), .data[[to_plot]],"genus",log1p)
p2 <- functaxboxplot(D_sub, (family),.data[[to_plot]], "family",log1p)
p3 <- functaxboxplot(D_sub, (order), .data[[to_plot]],"order",log1p)
p4 <- functaxboxplot(D_sub, (class),.data[[to_plot]], "class",log1p)
p5 <- functaxboxplot(D_sub, (phylum),.data[[to_plot]], "phylum",log1p)

plot_grid(p0,p1, p2, p3, p4,p5,labels ="auto")

```

```{r}

# Lets try for all different n16

D_sub <- mutate(D, n16 = factor(as.integer(n16)))
ggplot(D_sub) +
    geom_histogram(aes(x=log1p(div), y=after_stat(density))) +
  facet_wrap(~n16)


```
1 n16 always has zero, ofc. It seems the more n16 we get the less with zero div. Which again makes sense


Furthermore i believe im going to continue working from a speceis level as the distributions look ok at that level and it is not worth sacreficing resolution to get better model assumptions.. for now

### Div v n16
We expect there to be a relationshio.. lets see


## Antibiotics
Lets start by checking in general
The hypthesis was that they might have more div. 

```{r}
p1 <- ggplot(D)+
  geom_boxplot(aes(x=antibiotics, y=log1p(div)))
p2 <- ggplot(D)+
  geom_boxplot(aes(x=antibiotics, y=log(div)))
p3 <- ggplot(D)+
  geom_boxplot(aes(x=antibiotics, y=log(n16)))
p4 <- ggplot(D)+
  geom_boxplot(aes(x=antibiotics, y=(n16)))
plot_grid(p1,p2,p3,p4,labels="auto")
```
### Formatting daTA
Lets check for more specifc types of AR
First getting the subset of the data with AR resistence info about the Antibiotics
which target the 16s rRNA
```{r}
# Getting the ones which are actually targeting 16S 
# Reading them from ARtarget16s.csv
target16S <- read_csv2("../data/ARtarget16s.csv",show_col_types = FALSE,col_names = FALSE)
targetvector <- as.array(target16S$X1)
found_16S <- as.array(colnames(select(D_tmp,lincomycin:spiramycin.II)))
intersect <- intersect(targetvector,found_16S)
D_ar <- select(D_tmp, intersect, n16, div)

```
### Different types
#### Div
Now lets look at some plots
firstly for div
```{r}
library(gridExtra)

plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(D_ar)+
    geom_boxplot(aes(x=.data[[antibiotic]], y=log1p(div)))
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)
```


#### n16
and now for n16
```{r}
plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(D_ar)+
    geom_boxplot(aes(x=.data[[antibiotic]], y=log(n16)))
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)
```
### Deep dive into topramycin
Some seem to have a change but it is my suspesion that its because of the randomness with a small subset randomly having more or just because of other factors. Eg the small subset which are Antibiotic resistenet to a specific AB might just share taxonmic. Lets check for tobramycin
```{r}
# Get the taxonomy info for all the different bacteria resistent to tobramycin
D_ar_tax <- select(D_tmp, intersect, n16, div,genus,family,order, class,phylum)
filter(D_ar_tax, tobramycin == "R") %>% 
  select(genus,family,order,class,phylum) 
# It seems most of them share the Gammaproteobacteria class / are from the..


GorNot <- mutate(D_ar_tax, isGamma = ifelse(class == "Gammaproteobacteria","Gamma","notGamma")) %>% 
  mutate(isGamma=factor(isGamma)) 

p1 <- ggplot(GorNot) +
geom_boxplot(aes(x=isGamma, y=log1p(div)))

p2 <- ggplot(D_ar)+
    geom_boxplot(aes(x=tobramycin, y=log1p(div)))
                 

p3 <- ggplot(GorNot) +
geom_boxplot(aes(x=isGamma, y=log(n16)))

p4 <- ggplot(D_ar)+
    geom_boxplot(aes(x=tobramycin, y=log(n16)))
          
plot_grid(p1,p2,p3,p4)


```

So we can see that it's proberly just tax- information messing it up
Statistically we should test types of AR agains tax.
Here we have 895 different in Gammaproteo group but only 8 of them with the AR

```{r}
#Lastly lets try and plot the ones with AR against the ones which do not for gammaproteo
# Its mostly saying for the ones which are Gammaproteo

GorNot %>% 
  ggplot()+
    geom_boxplot(aes(x=tobramycin, y=log1p(div))) +
    facet_wrap(~isGamma)

# There actually seem to be more
# Lest see for n16
# We expect n16 to go up for R again since the are correlated
GorNot %>% 
  ggplot()+
    geom_boxplot(aes(x=tobramycin, y=(n16))) +
    facet_wrap(~isGamma)
# It seems to actually go down ! 
# so there is something going on :I

```

Lets try and remove the tax (maybe phylum (could be class)) information

```{r}
library(modelr)
D_ar_tax %>% 
  ggplot() +
  geom_boxplot(aes(x=phylum, y=log1p(div)))

D_ar_tax %<>% mutate(phylum = factor(phylum)) 
# Fit the model
tax_model <- lm(log1p(div) ~phylum ,data=D_ar_tax)
summary(tax_model)

# Visualize the predictions
pred <- D_ar_tax %>% 
  data_grid(phylum) %>% 
  add_predictions(tax_model)
ggplot() +
  geom_boxplot(aes(y=phylum, x=log1p(div)), data=D_ar_tax) +
  geom_point(aes(y=phylum, x=pred), data=pred, col="red")

# Visulize the residuals
res <- D_ar_tax %>% 
  add_residuals(tax_model)
# no need to specify x,y in aes()
# We can see the effect here
res %>% 
  ggplot(aes(y=resid, x=tobramycin)) +
  geom_boxplot()

# Looking at the res there is still a lot of uneqplanied especially under
res %>% 
  ggplot(aes(resid)) +
  geom_histogram()
D_tmp
# Lets see where it comes from
res %>% 
  ggplot(aes(n16 ,resid)) +
  geom_point(size=0.5) +
  geom_ref_line(h=0, col = "red", size = 1)



```

```{r}
library(magrittr)
# trying to take into consideration the n16
tax_n16_model <- update(tax_model, ~.+n16)
# Now it looks alot better:
D_ar_tax %>% 
  add_residuals(tax_n16_model) %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res <- D_ar_tax %>% 
  add_residuals(tax_n16_model) 
  
plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(res)+
    geom_boxplot(aes(x=.data[[antibiotic]], y=resid))
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)


```

## Lets see for class
```{r}

# making the lm model
tax_n16_model <- lm(log1p(div) ~ class+n16, D_ar_tax)
# This looks better than before. Maybe go all the way down?
D_ar_tax %>% 
  add_residuals(tax_n16_model) %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res <- D_ar_tax %>% 
  add_residuals(tax_n16_model) 
  
plotlist = list()
for(i in seq_along(intersect)){
  antibiotic = intersect[i]
  p <- ggplot(res)+
    geom_boxplot(aes(x=.data[[antibiotic]], y=resid))
  plotlist = c(plotlist, list(p))
}

wrap_plots(plotlist)


res %>% 
  ggplot(aes(log(n16) ,resid)) +
  geom_point(size=0.5) +
  geom_ref_line(h=0, col = "red", size = 1) +
  geom_smooth()

```

# Lets try with the log(n16)
```{r}

D_ar_sub <- mutate(D_ar_tax,logdiv = log1p(div), logn16 = log(n16))

# making the lm model
tax_n16_model <- lm(logdiv ~ phylum + logn16, data=D_ar_sub)

# This looks better than before. Maybe go all the way down taxonomically?
D_ar_sub %>% 
  add_residuals(tax_n16_model) %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res <- D_ar_sub %>% 
  add_residuals(tax_n16_model) 
  

#plotlist = list()
#for(i in seq_along(intersect)){
#  antibiotic = intersect[i]
#  p <- ggplot(res)+
#    geom_boxplot(aes(x=.data[[antibiotic]], y=resid))
#  plotlist = c(plotlist, list(p))
#}
#
#wrap_plots(plotlist)


res
res %>% 
  ggplot(aes(2^logn16 ,resid)) +
  geom_hex(bins=50) +
  geom_ref_line(h=0, col = "red", size = 1) +
  geom_smooth()
# Here it looks good, might be because some genera only have like two species in them..

```

FOR 
```{r}
D_ar_sub <- mutate(D_ar_tax,logdiv = log1p(div), logn16 = log(n16))

# making the lm model
tax_n16_model <- lm(logdiv ~ order + logn16, data=D_ar_sub)

# This looks better than before. Maybe go all the way down taxonomically?
D_ar_sub %>% 
  add_residuals(tax_n16_model) %>% 
  ggplot(aes(resid)) + geom_histogram(bins = 40,binwidth = 0.1) 

res <- D_ar_sub %>% 
  add_residuals(tax_n16_model) 
  

#plotlist = list()
#for(i in seq_along(intersect)){
#  antibiotic = intersect[i]
#  p <- ggplot(res)+
#    geom_boxplot(aes(x=.data[[antibiotic]], y=resid))
#  plotlist = c(plotlist, list(p))
#}
#
#wrap_plots(plotlist)


res
res %>% 
  ggplot(aes(2^logn16 ,resid)) +
  geom_hex(bins=50) +
  geom_ref_line(h=0, col = "red", size = 1) +
  geom_smooth()
# Here it looks good, might be because some ORDERs only have like two species in them..


# Lets look at order distribution
D_ar_tax %>% 
  group_by(phylum) %>% 
  summarise(n = n()) %>% 
  mutate(tax_ordered = fct_reorder(phylum,desc(n))) %>% 
  filter(n < 10) %>% 
  ggplot() +
  geom_segment(aes(x=tax_ordered, xend = tax_ordered,y=0, yend=n))

s

# 10 out of 34 tax have under 10 entries. Largest Pseudomonadota have 1800
D_ar_tax %>% 
  group_by(phylum) %>% 
  summarise(n = n()) %>% 
  mutate(tax_ordered = fct_reorder(phylum,desc(n))) %>% 
  filter(n < 10) %>% 
  nrow()


D_ar_tax %>% 
  group_by(phylum) %>% 
  summarise(n = n()) %>% 
  mutate(tax_ordered = fct_reorder(phylum,desc(n))) %>% 
  nrow()


D_ar_tax %>% 
  group_by(phylum) %>% 
  summarise(n = n()) %>% 
  mutate(tax_ordered = fct_reorder(phylum,desc(n))) %>% 
  arrange(desc(n))

nrow(D_tmp)
```
Probelemet ift tax er at det for de små tax basically bare giver dem svaret




## Genome size v total seq length

```{r}
ggplot(D, aes(x=total_seq_length, y=n16, label=species)) + 
  geom_point()


```



